---
title: "Linear Regression"
author: "Akarin Luckanasub"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### **Introduction**
##### This document demonstrates how to perform both simple and multiple linear regression on the mtcars dataset using the dplyr library for data manipulation. The goal is to predict the miles per gallon (mpg) based on various predictor variables.

## **Simple Linear Regression**
##### Start with a simple linear regression where mpg is predicted based on hp (horsepower).

### **Fit the Model**

```{r}
# Simple linear regression
lmfit <- lm(mpg ~ hp, data = mtcars)

# Summary of the model
summary(lmfit)
```

### **Predictions**
##### Make predictions using the fitted model.

```{r}
# Prediction with a specific horsepower value
predicted_mpg_200hp <- lmfit$coefficients[[1]] + lmfit$coefficients[[2]] * 200
predicted_mpg_200hp

# Predict for multiple values
newcar <- data.frame(
  hp = c(250, 320, 400, 410, 450)
)

# Predict mpg for the new values
newcar$mpg_predicted <- predict(lmfit, newdata = newcar)

# Display the predictions
newcar
```

## **Multiple Linear Regression**
##### Extend the model to include additional predictor variables: wt (weight) and am (transmission).

### **Fit the Model**

```{r}
# Multiple linear regression
lmfitv2 <- lm(mpg ~ hp + wt + am, data = mtcars)

# Coefficients
coefs <- coef(lmfitv2)
coefs

```

### **Full Model**
##### build a full model using all predictors except gear.

```{r}
# Full model with all predictors except gear
lmfitfull <- lm(mpg ~ . - gear, data = mtcars)

# Predictions using the full model
mtcars$predicted <- predict(lmfitfull)
```

### **Calculate RMSE**
##### Calculate the Root Mean Squared Error (RMSE) for the model.

```{r}
# Squared error
sq_error <- (mtcars$mpg - mtcars$predicted) ^ 2

# RMSE
rmse <- sqrt(mean(sq_error))
rmse
```

## **Full Modelling**
##### Split the data into training and test sets, train the model, and evaluate its performance.

### **Split the Data**

```{r}
# Set seed for reproducibility
set.seed(42)

# Number of rows in the dataset
n <- nrow(mtcars)

# Generate random indices for training data (80% of the data)
id <- sample(1:n, size = n * 0.8)

# Create training and test datasets
train_data <- mtcars[id, ]
test_data <- mtcars[-id, ]

```

### **Train the Model**

```{r}
# Train the model on training data
model1 <- lm(mpg ~ hp + wt + am + disp, data = train_data)

# Predictions on training data
predicted_train_model1 <- predict(model1)

# Calculate training error
error_train_model1 <- train_data$mpg - predicted_train_model1

# Calculate RMSE for training data
rmse_train_model1 <- sqrt(mean((error_train_model1) ^ 2))
rmse_train_model1

```

### **Test the Model**

```{r}
# Predictions on test data
predicted_test_model1 <- predict(model1, newdata = test_data)

# Calculate test error
error_test_model1 <- test_data$mpg - predicted_test_model1

# Calculate RMSE for test data
rmse_test_model1 <- sqrt(mean((error_test_model1) ^ 2))
rmse_test_model1

```

### **Print Results**

```{r}
# Print RMSE results
cat("RMSE Train:", rmse_train_model1, "\n")
cat("RMSE Test:", rmse_test_model1)

```

### **Conclusion**
##### This document demonstrated the process of performing both simple and multiple linear regression on the mtcars dataset. The RMSE values for the training and test datasets provide a measure of the model's performance.



